---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# datenanpassr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/datenanpassr)](https://CRAN.R-project.org/package=datenanpassr)
<!-- badges: end -->

The goal of datenanpassr is to ...

## Installation

You can install the released version of datenanpassr from [CRAN](https://CRAN.R-project.org) with:

```{r, eval=FALSE}
devtools::install_gitlab("urswilke/datenanpassr")
```

DOESN'T WORK YET BECAUSE REPO IS PRIVATE!!!

## Example

### Included labelled data

The package has an included dataset `fake_survey` with made-up data of a survey. The data is also included as an SPSS sav file [fake_survey.sav](inst/extdata/fake_survey.sav). These are generated by [this Rmd file](devel/make_fake_data.md). After installing the package, here is how you can access these data sets: 

```{r libs, message=FALSE}
library(datenanpassr)
# Needed for the printing of the labels when printing the dataframe:
library(tidyverse)
```


```{r example}
datenanpassr::fake_survey
path <- system.file("extdata", "fake_survey.sav", package = "datenanpassr")
haven::read_sav(path)
```

### Apply data adjustments on labelled data

Suppose you have an SPSS data file 

```{r}
spss_filepath <- system.file("extdata", "fake_survey.sav", package = "datenanpassr")
df <- haven::read_sav(spss_filepath)
df
```


and want to modify some of the content. 

In datenanpassr there's, an [Excel file](inst/extdata/mapping.xlsx) included to demonstrate how this can be done with this package:

```{r}
mapping_filepath <- system.file("extdata", "mapping.xlsx", package = "datenanpassr")
df_mod <- mapp_xl_to_data(fake_survey, mapping_filepath)
df_mod
```

The maipulations defined in the Excel file are applied on the dataframe of the SPSS package.

You can save this dataframe back to an SPSS file by using the [haven package](https://haven.tidyverse.org/):

```{r, eval=FALSE}
haven::write_sav(df_mod, "fake_survey_mod.sav")
```

