---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# datenanpassr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/datenanpassr)](https://CRAN.R-project.org/package=datenanpassr)
<!-- badges: end -->



The R package datenanpassr is an approach to programmatically manipulate labelled datasets via a pre-defined syntax of various types of commands in various types of Excel sheets. It is a replacement of what my brother initially programmed in VBA and SPSS and how we approach our daily work of data cleaning of survey data. The package can be used to write various data manipulations in a concise way filling in commands in Excel cells.



## Installation

You can install datenanpassr from Gitlab with:

```{r, eval=FALSE}
devtools::install_gitlab("urswilke/datenanpassr")
```

## Example

First load the library:

```{r}
library(datenanpassr)
```

### Apply data adjustments on labelled data

Suppose you have an SPSS data file 

```{r}
spss_file <- system.file("extdata", "fake_survey.sav", package = "datenanpassr")
df <- haven::read_sav(spss_file)
df
```


and want to modify some of the content. 

<!-- TODO:  -->

<!-- You can create an Excel mapping file that's based on a template filled with variable and label information from the dataset: -->


<!-- ```{r, eval=FALSE} -->
<!-- datenanpassr::mapp_create(df, "mapping.xlsx") -->
<!-- ``` -->

In the package you can find an [example Excel mapping file](inst/extdata/mapping.xlsx) to demonstrate the commands in this package. If you install `datenanpassr`, you can access the path of this file with

```{r}
mapping_file <- system.file("extdata", "mapping.xlsx", package = "datenanpassr")
```

and then open it with:
```{r, eval=FALSE}
utils::browseURL(mapping_file)
```


Have a look at the Excel command block vignette for examples how to manipulate and generate new variables in your labelled dataset:

```{r, eval=FALSE}
vignette("excel_command_blocks")
```

There you can find out more about the syntax how the commands in the mapping file work.


Once you have added the commands to manipulate your data, you can generate a mapping object with


```{r, message=FALSE, warning=FALSE}
mapping <- Mapping$new(df, mapping_file)
```

and apply the changes to your dataset with:

```{r}
mapping$modify_data()
```
You can  then access the modified data with:
```{r}
(df_mod <- mapping$dat_mod)
```


Let's also have a closer look at one of the new variables:

```{r}
df_mod$q6n
```


The manipulations defined in the Excel files are applied on the dataframe that was derived from the SPSS file. 

You can save the dataframe back to an SPSS file by again using the [haven package](https://haven.tidyverse.org/):

```{r, eval=FALSE}
haven::write_sav(df_mod, "fake_survey_mod.sav")
```

