---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# datenanpassr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/datenanpassr)](https://CRAN.R-project.org/package=datenanpassr)
<!-- badges: end -->

The goal of datenanpassr is to ...

## Installation

You can install the released version of datenanpassr from [CRAN](https://CRAN.R-project.org) with:

```{r, eval=FALSE}
devtools::install_gitlab("urswilke/datenanpassr")
```

DOESN'T WORK YET BECAUSE REPO IS PRIVATE!!!

## Example

### Apply data adjustments on labelled data

Suppose you have an SPSS data file 

```{r}
spss_filepath <- system.file("extdata", "fake_survey.sav", package = "datenanpassr")
df <- haven::read_sav(spss_filepath)
df
```


and want to modify some of the content. 

You can create an Excel mapping file that's based on a template filled with variable and label information from the dataset:


```{r, eval=FALSE}
mapp_create(df, "mapping.xlsx")
```

In this Excel file commands can be added to do data cleaning or create new variables in the dataset.

In datenanpassr there's, an [Excel file](inst/extdata/mapping.xlsx) included to demonstrate how this can be done with this package. Once you have added the commands to manipulate your data, you can apply the changes to your dataset with:

```{r, message=FALSE, warning=FALSE}
mapping_filepath <- system.file("extdata", "mapping.xlsx", package = "datenanpassr")
df_mod <- mapp_xl_to_data(df, mapping_filepath)
```


```{r}
df_mod
```
Let's also have a closer look at one of the new variables:

```{r}
df_mod$q6n
```


The manipulations defined in the Excel files are applied on the dataframe that was derived from the SPSS file. Have a lock at the Excel command block vignette for examples how to manipulate and generate new variables in your labelled dataset:

```{r, eval=FALSE}
browseVignettes("datenanpassr")
```


You can save the dataframe back to an SPSS file by again using the [haven package](https://haven.tidyverse.org/):

```{r, eval=FALSE}
haven::write_sav(df_mod, "fake_survey_mod.sav")
```

