---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# datadaptor

*The main home of this repository is [here](https://codeberg.org/urswilke/datadaptor)*

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/datadaptor)](https://CRAN.R-project.org/package=datadaptor)
[![datadaptor status badge](https://urswilke.r-universe.dev/datadaptor/badges/version)](https://urswilke.r-universe.dev/datadaptor)
<!-- badges: end -->



The R package datadaptor is an approach to programmatically manipulate labelled datasets via a pre-defined syntax of various types of commands in various types of Excel sheets. It is a replacement of what my brother initially programmed in VBA and SPSS and how we approach our daily work of data cleaning of survey data. The package can be used to write various data manipulations in a concise way filling in commands in Excel cells.



## Installation

You can install datadaptor from codeberg with:

```{r, eval=FALSE}
devtools::install_git("https://codeberg.org/urswilke/datadaptor")
```

Otherwise, you can also install datadaptor from github or gitlab with:

```{r, eval=FALSE}
devtools::install_github("urswilke/datadaptor")
devtools::install_gitlab("urswilke/datadaptor")
```

## Example

First load the library:

```{r}
library(datadaptor)
```

### Apply data adjustments on labelled data

Suppose you have an SPSS data file 

```{r}
spss_file <- system.file("extdata", "fruit_survey.sav", package = "datadaptor")
df <- haven::read_sav(spss_file)
df
```


and want to modify some of the content. 

<!-- TODO:  -->

<!-- You can create an Excel mapping file that's based on a template filled with variable and label information from the dataset: -->


<!-- ```{r, eval=FALSE} -->
<!-- datadaptor::create_mapping(df, "mapping.xlsx") -->
<!-- ``` -->

In the package you can find an [example Excel mapping file](https://codeberg.org/urswilke/datadaptor/src/branch/master/inst/extdata/mapping-fruits.xlsx) to demonstrate the commands in this package. If you install `datadaptor`, you can access the path of this file with

```{r}
mapping_file <- system.file("extdata", "mapping-fruits.xlsx", package = "datadaptor")
```

and then open it with:
```{r, eval=FALSE}
utils::browseURL(mapping_file)
```


Have a look at the `vignette("command_blocks")` for examples how to manipulate and generate new variables in your labelled dataset.
There you can find out more about the syntax how the commands in the mapping file work.
Once you have added the commands to manipulate your data, you can generate a mapping object with


```{r, message=FALSE, warning=FALSE}
mapping <- Mapping$new(df, mapping_file)
```

and apply the changes to your dataset with:

```{r}
mapping$modify_data()
```
You can  then access the modified data with:
```{r}
(df_mod <- mapping$dat_mod)
```

Let's also have a closer look at one of the new variables:

```{r}
df_mod$kq5
```


The manipulations defined in the Excel file are applied on the dataframe that was derived from the SPSS file. 

You can save the dataframe back to an SPSS file by again using the [haven package](https://haven.tidyverse.org/):

```{r, eval=FALSE}
haven::write_sav(df_mod, "fruit_survey-modified.sav")
```

[Here](https://urswilke.codeberg.page/crosstabser/articles/crosstabser.html#mapping-file-content) 
you can find a small description what each of the commands in `mapping_file` does,
and how this can be used for the tabulation of the data.


## Further use

### Tabulation

If you additionally want to tabulate survey data, please have a look at 
the accompanying package [crosstabser](https://codeberg.org/urswilke/crosstabser/)
that expands the functionality of datadaptor.

### Dashboards

For interactive dashboards with charts of the crosstabs, please have a look at
the accompanying package [table_charter](https://gitlab.com/urswilke/table_charter)
that expands the functionality of crosstabser.

### datadaptor - crosstabser - table_charter

For an interactive demo of the combined use 
of datadaptor, crosstabser & table_charter, 
where you can play around in your browser and see how that changes the results,
please have a look 
[here](https://urswilke.github.io/datadaptor-crosstabser-table_charter-demo).

## Security warning

datadaptor evaluates code from user input. 
Do not expose this program to the internet or random users under any circumstances.
