---
title: "Documentation of the Excel mapping command blocks"
output: 
  rmarkdown::html_vignette:
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Documentation of the Excel mapping command blocks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE
)
library(datenanpassr)
library(dplyr)
```

```{css css_def}
.parent {
  display: flex;
  height: 300px; 
}

.child {
  width: 100px; 
  height: 100px; 
  margin: auto; 
}


.flex-container {
  display: flex;
  align-items: center;
  flex-flow: row wrap;
  column-gap: 20px;
  row-gap: 10px;
}
```

```{r gt_style, child="../man/fragments/gt_table_stylers.Rmd", include=FALSE}
```
# Command block overview

In this document, we will show how the different command blocks in datenanpassr work. 
There are the following keywords that you can use to enter the command blocks in your Excel mapping file:
```{r cdb_classes}
command_block_classes |> excelish()
```

*The column* `"command_block"` *shows the class, which determines how 
datenanpassr parses the command block* (with `parse_command_args()`) *and 
then applies it on the data* (with `apply_command()`).

# Command block examples

```{r fun_defs_illu, echo=FALSE}
process_cdb_raw <- function(dat, cmd_block_tibble, sheet_cat, ...) {
  mapping_file <- gen_mapping_obj(cmd_block_tibble, sheet_cat)
  Mapping$new(dat, mapping_file, ...)$modify_data()$dat_mod
}
gen_mapping_obj <- function(cmd_block_tibble, sheet_cat) {
  if (sheet_cat == "Free") {
    cmd_block_tibble <- bind_rows(
      tibble(
        X1 = character(),
        X2 = character(),
        X3 = character(),
        X4 = character(),
        X5 = character(),
      ),
      cmd_block_tibble |> mutate(row = row_number())
    )
  }
  if (sheet_cat == "Variables") {
    cmd_block_tibble <- bind_rows(
      tibble(
        var = character(),
        op = character(),
        new_name = character(),
        new_label = character(),
      ),
      cmd_block_tibble |> mutate(row = row_number())
    )
  } 
  if (sheet_cat == "Label") {
    cmd_block_tibble <- bind_rows(
      tibble(
        var = character(),
        nv = character(),
        new_label = character(),
        sum_var_label = character(),
        sum_var_value = character(),
        sum_var_vallab = character(),
      ),
      cmd_block_tibble |> mutate(row = row_number())
    )
  } 
  tibble::lst(!!sheet_cat := cmd_block_tibble)
}

tab_labels_wide <- function(df) {
  varl_table <- df |>
    datenanpassr:::tab_varlabs(remove_empty = FALSE) |> 
    tidyr::pivot_wider(names_from = var, values_from = varlab) |> 
    dplyr::mutate(val = "varlab", .before = 1)
  vall_table <- df |>
    datenanpassr:::tab_vallabs(remove_empty = FALSE) |>
    dplyr::rename(val = nv) |> 
    tidyr::pivot_wider(names_from = var, values_from = vallab) |> 
    dplyr::mutate(val = as.character(val))
  labs <- dplyr::bind_rows(varl_table, vall_table)
  if (nrow(labs) == 0) {
    labs <- labs |> dplyr::bind_rows(dplyr::tibble(val = NA_character_))
  }
  labs
}

join_labs <- function(df) {
  labs <- tab_labels_wide(df)
  df |>
    dplyr::mutate(
      dplyr::across(
        .cols = everything(),
        .fns = as.character
      ),
      val = NA_character_,
      .before = 1,
      type = "values"
    ) |> 
    dplyr::bind_rows(
      labs |> dplyr::mutate(type = "labels")
    )
}


gen_data_gt_tables <- function(dat, cmd_block_tibble, sheet_cat, ...) {
  dat_mod <- process_cdb_raw(dat, cmd_block_tibble, sheet_cat, ...)
  
  dat_table <- join_labs(dat)
  dat_mod_table <- join_labs(dat_mod)
  
  dat_gt <- dat_table |> 
    format_my_gt()
  dat_mod_gt <- dat_mod_table |>
    format_my_gt()
  
  tibble::lst(
    dat_gt,
    dat_mod_gt
  )

}
format_my_gt <- function(gt_obj) {
  gt_obj |>
    gt::gt(
      groupname_col = "type"
    ) |>
    gt::tab_style(
      style = list(
        gt::cell_fill(color = "#D4E6B4")
      ),
      locations = gt::cells_body(
        rows = val == "varlab")
    ) |> 
    gt::tab_style(
      style = list(
        gt::cell_fill(color = "#D7E9E9")
      ),
      locations = gt::cells_body(rows = type == "labels" & val != "varlab")
    ) |> 
    gt::tab_style(
      style = list(
        gt::cell_text(color = "#999999")
      ),
      locations = gt::cells_body(rows = type == "labels")
    ) |> 
    gt::sub_missing(missing_text = "") |> 
    gt::cols_label(val = "") |> 
    gt::opt_table_lines() |> 
    gtExtras::gt_add_divider(columns = "val", style = "dashed")
}

rename_cols_for_vignette <- function(df, sheet_cat) {
  if (sheet_cat == "Free") {
    df <- purrr::set_names(df, LETTERS[seq_len(ncol(df))]) 
  }
  df
}

illustrate_cmd_block <- function(dat, cmd_block_tibble, sheet_cat, ...) {
  cmd_block_gt <- cmd_block_tibble |>
    rename_cols_for_vignette(sheet_cat) |>
    excelish()
  
  gts <- gen_data_gt_tables(dat, cmd_block_tibble, sheet_cat, ...)
  htmltools::div(
    class = "flex-container", 
    htmltools::div(style = "align-self: flex-start", gts$dat_gt),
    htmltools::div(style = "font-size: 45px", "&"),
    htmltools::div(style = "flex-item", cmd_block_gt),
    htmltools::div(style = "font-size: 45px", "=>"),
    htmltools::div(style = "align-self: flex-start", gts$dat_mod_gt),
  )
  # This generates the following html:
  # <div class="flex-container">
  #   <div style="align-self: flex-start">`r gts$dat_gt`</div>
  #   <div style="font-size: 45px">&</div>
  #   <div class="flex-item">`r cmd_block_gt`</div>
  #   <div style="font-size: 45px">=></div>
  #   <div style="align-self: flex-start">`r gts$dat_mod_gt`</div>
  # </div>
  
  # if you want to see the result of this function interactively, pipe it in the following function:
  # view_html <- function(x){
  #   html <- paste(x, collapse = '\n')
  #   tf <- tempfile(fileext = ".html")
  #   writeLines(html, tf)
  #   rstudioapi::viewer(tf)
  # }
}

```

# `Free` sheets

Here we'll look at examples how various commands in the `Free` sheets can look
like.

```{r, echo=FALSE}
sheet_cat <- "Free"
```


## `#VARL`

```{r varl}
dat <- tibble(q1 = 4:6)

cmd_block_tibble <- tibble(X1 = "#VARL", X2 = "q1", X3 = "assigned variable label")

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```


Actually, this command does not change the values, but it assigns the variable 
label specified in column **C** of the command to the variable specified in 
column **B**.



## `#VALL`

Assign **val**ue **l**abels

```{r vall}
dat <- tibble(q1 = c(1, 1, 2))

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3,
  "#VALL", "q1", "yes/no question",
   NA, "1", "Yes",
   NA, "2", "No"
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

This command also does not change the values. Just like `#VARL`, it assigns the variable label
specified in the first row of column **C** (if empty, it keeps the variable
label, if already existing) to the variable specified in the first row in column
**B**. It also assigns the value labels in column **C** to 
the values in column **B** in the following rows of the command.


## `#AVALL`

**A**dd **val**ue **l**abels

```{r avall}
dat <- tibble(q1 = haven::labelled(c(1, 2, 3), c(a = 1)))

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3,
  "#AVALL", "q1", NA_character_,
   NA, "2", "b",
   NA, "3", "c"
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

`#AVALL` does the same as `#VALL`, except that it keeps the value values of the 
original variable, if they are not overwritten.

## `#RMVAL`

**R**e**m**ove **val**ues and their labels

```{r rmval}
dat <- tibble(q1 = haven::labelled(c(1, 2, 3), c(a = 1, b = 2, c = 3), "abc"))

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3, ~X4,
  "#RMVAL", "q1", "q2", "def",
   NA, "2", "", NA_character_,
   NA, "3", "", NA_character_,
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

Here, the `#RMVAL` command makes a copy of `q1`, named `q2`, and removes the 
specified values.
(In the first line of the command block, column **B** specifies the variable to be recoded, 
and in column **C** the name of the recoded variable. Column **D** assigns the new 
variable label.) In the following rows you specify the values and their labels 
to be removed in column **B**.

## `#COMP`

### Example

```{r comp1}
dat <- tibble(q1 = 1:3)

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3,
  "#COMP", "q2", "5",
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```
This command **comp**utes a new variable (specified in column **B**) with the value 
in column **C**.

### Example

You can also assign a variable to the values of the new variable. In this case,
the labels of the original variable are also copied:
```{r comp2}
dat <- tibble(q1 = haven::labelled(c(1, 2, 3), c(a = 1, b = 2, c = 3), "abc"))

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3,
  "#COMP", "q2", "q1",
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

## `#IF`

### Example

```{r if1}
dat <- tibble(q1 = 1:3)

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3,
  "#IF", "q1 >= 2", "q2 = 5",
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```
This command takes a condition (more precisely a logical vector of length of the 
variables in the data) in column **B** and an assignment in column **C** (of the form 
`<variable name> = <assigned values>`). 
When the condition is `TRUE`, the expression (`<assigned values>`) is assigned 
to `<variable name>`.
In this case, the value 5 is assigned to `q2` where `q1` is greater or equal to 2.
In this case `q2` didn't exist in the data before, and thus is initialized to
`NA` before.

### Example

If the variable existed before, the values where the condition evaluates to `FALSE`
are kept, as can be seen in the next example:

```{r if2}
dat <- tibble(q1 = 1:3, q2 = c(7, 8, 9))

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3,
  "#IF", "q1 >= 2", "q2 = 5",
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

Here the value of 7 in the first row is kept from the original data.

### Example

The `#IF` command preserves labels, if an already existing variable is modified:

```{r if3}
dat <- tibble(q1 = haven::labelled(c(1, 2, 3), c(a = 1, b = 2, c = 3), "abc"))

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3,
  "#IF", "q1 >= 2", "q1 = 5",
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

## #REC

### Example

```{r rec1}
dat <- tibble(q1 = haven::labelled(c(1, 2, 3, 4, 5, 99), c(a = 1, b = 2, c = 3, d = 4, e = 5, "No answer" = 99), "abcde"))

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3, ~X4, ~X5,
  "#REC", "q1", "kq1", "fgh", NA_character_,
  "", "1", "2", "1", "a & b",
  "", "3", "", "2", "c",
  "", "4", "5", "3", "d & e",
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

The `#REC` command **rec**odes variables. In this example `q1` is recoded to `q2`.
(In the first line of the command block, column **B** specifies the variable to be recoded, 
and in column **C** the name of the recoded variable. Column **D** assigns the new 
variable label.) In the following rows you specify the intervals to be recoded 
in column **B** and **C**, column **D** specifies the new value and column **E**
its label. So for instance, the second line specifies, that all values in `q1` 
between 1 and 2 are recoded to 1 in `q2` with the value label `a & b`.

### Example

If you don't specify the variable label (column **D**, first row), the one of 
the original variable is copied:

```{r rec2}
dat <- tibble(q1 = haven::labelled(c(1, 2, 3), c(a = 1, b = 2, c = 3), "abc"))

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3, ~X4, ~X5,
  "#REC", "q1", "kq1", NA_character_, NA_character_,
  "", "1", "2", "1", "a & b",
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

### Example

If the name of the recoded variable is the same as the original variable, the 
values that are not recoded are kept:
```{r rec3}
dat <- tibble(q1 = haven::labelled(c(1, 2, 3), c(a = 1, b = 2, c = 3), "abc"))

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3, ~X4, ~X5,
  "#REC", "q1", "q1", NA_character_, NA_character_,
  "", "1", "2", "0", "a & b",
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

## #KG

### Example

```{r kg1}
dat <- tibble(
  q1 = haven::labelled(
    c(1, 2, 1, 2, 1, 2), 
    c(good = 1, bad = 2),
    "How are you?"
  ),
  q2 = haven::labelled(
    c(1, 2, 3, 1, 2, 3), 
    c(UK = 1, US = 2, FR = 3, DE = 4),
    "Country"
  ),
)

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3, 
  "#KG", "q1", "q2", 
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

The `#KG` command is useful to calculate cross-tabulations for subsets. It 
computes new variables for all the value labels of the variable `q2` (in column **C**).
The value labels are copied from `q1` (the variable in column **B**). The new variable 
labels are a combination of the value labels of `q2` and the variable label of `q1`.
The values are copied of `q1` if `q2` takes the value specified in the variable 
label and are `NA` otherwise.

## #SELECT

This command block passes the arguments in column **B** to `dplyr::select()`. However, renaming 
doesn't work (see `#RENAME` to do this). It can be used to remove and reorder 
variables in the dataset.

### Example

This removes the variable `a2`:

```{r select1}
dat <- tibble(
  a1 = c(1, 2),
  b1 = c(3, 4),
  b2 = c(1, 2),
  a2 = c(5, 6),
)

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, 
  "#SELECT", "-a2"
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

### Example

This only selects the variables matching the regular expression in the quotes
(variables containing the letter a followed by a digit):

```{r select2}
dat <- tibble(
  a1 = c(1, 2),
  b1 = c(3, 4),
  b2 = c(1, 2),
  a2 = c(5, 6),
)

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, 
  "#SELECT", 'matches("a\\\\d")'
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

### Example
In order to keep all variables, and only reorder those named before, you can 
pass `everything()` in the end:

```{r select3}
dat <- tibble(
  a1 = c(1, 2),
  b1 = c(3, 4),
  b2 = c(1, 2),
  a2 = c(5, 6),
)

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, 
  "#SELECT", 'matches("a\\\\d")',
  "", 'everything()',
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```


## #RENAME

### Example

This command renames variables in column **B** to the names in column **C**:

```{r rename1}
dat <- tibble(
  a1 = c(1, 2),
  b1 = c(3, 4),
  b2 = c(1, 2),
  a2 = c(5, 6),
)

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3,
  "#RENAME", 'a2', "new_name_for_a1",
  "", 'b2', "new_name_for_b2"
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

# `curlychop()`

My brother originally called this function `severalize()`. As I was having a 
hard time getting to used to this work back then, I decided that I wanted to 
give it another name that at least sounds more interesting. I ended up with 
`curlychop()` (click link for the documentation of the underlying R function). 

The idea is to prevent you from writing redundant code by turning one line into
several commands. Let's have a first look 
at it with the help of an example.

### Example

Let's assume you want to make copies of several variables, all ending with the
same suffix "1". You can do this by using several `#COMP` commands:

```{r curlychop1}
dat <- tibble(
  a1 = c(1, 2),
  b1 = c(3, 4),
)

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3,
  "#COMP", "ka1", "a1",
  "#COMP", "kb1", "b1",
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

The function `curlychop()` lets you write these commands in a more succint form:


```{r curlychop1_2}
tibble::tribble(
  ~X1, ~X2, ~X3,
  "#COMP", "k{a b}1", "{a b}1",
) |> excelish()
```

It replaces the curly braces by each of the parts inside (separated by spaces).
You can put as many parts inside the curly braces as you want, but in the examples
we'll only write 2 parts to save the eye.

### Example

You can also write multiple parts with curly braces inside one cell

```{r curlychop2}
dat <- tibble(
  a1 = c(1, 2),
  b1 = c(3, 4),
)

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3,
  "#IF", "{a b}1 == {1 4}", "k{a b}1 = {5 6}",
  )

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

The above command block is equivalent to these 2 `#IF` commands:

```{r curlychop_illustrator, echo=FALSE}
curlychop_illustrator <- function(df) {
  df |> 
    mutate(row = row_number()) |> 
    curlychop() |> 
    select(-row, -raw_index) |> 
    purrr::set_names(LETTERS[seq_len(ncol(df))]) |> 
    excelish()
}
```


```{r curlychop2_2}
cmd_block_tibble |> curlychop_illustrator()
```


### Example

For commands spanning multiple rows, you can also use `curlychop()` in the first 
row:

```{r curlychop3}
dat <- tibble(
  q_1 = haven::labelled(c(1, 2, 3), label = "varlab q1"),
  q_2 = haven::labelled(c(2, 3, 4), label = "varlab q2"),
)

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3, ~X4, ~X5,
  "#REC", "q_{1 2}", "kq{_1 _2}", NA_character_, NA_character_,
  NA, "1", "2", "1", "value label of 1",
  NA, "3", "4", "2", "value label of 2",
)

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

The above command block is equivalent to these 2 `#REC` commands:

```{r curlychop3_2}
cmd_block_tibble |> curlychop_illustrator()
```




# `Variables` sheets

To enter commands here, you don't need to specify the commands, as in the `Free` 
sheets. Instead you need to put the information in one of the columns `op`, 
`new_name` or `new_label`:

```{r, echo=FALSE}
sheet_cat <- "Variables"
```

## #RENAME_varsheet

### Example

This command renames the variables specified in the `var` column (column **A**)
to the names in the `new_name` column. It does the same as the `#RENAME` command.

```{r rename_varsheet}
dat <- tibble(
  a1 = c(1, 2),
  b1 = c(3, 4),
)

cmd_block_tibble <- tibble::tribble(
  ~var, ~new_name,
  "a1", 'c1', 
  "b1", 'd1', 
)

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

## #NEWLAB

### Example

This command assings new variable labels (specified in the `new_label` column)
to the variables specified in the `var` column (column **A**). It does the same
as the `#VARL` command.

```{r newlab}
dat <- tibble(
  a1 = haven::labelled(c(1, 2), label = "original variable label"),
  b1 = haven::labelled(c(1, 2), label = "original variable label of b1"),
  c1 = c(3, 4),
)

cmd_block_tibble <- tibble::tribble(
  ~var, ~new_label,
  "a1", "I'm new", 
  "b1", NA, 
  "c1", NA, 
)

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

## #DROP

### Example

This command removes the variables marked with `d` (for "drop") in the `op` column:

```{r drop}
dat <- tibble(
  a1 = c(1, 2),
  b1 = c(3, 4),
)

cmd_block_tibble <- tibble::tribble(
  ~var, ~op,
  "a1", 'd', 
  "b1", '', 
)

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

The result is the same as specifying `#SELECT -a1` on a `Free` sheet.

## #STR2NUM

### Example

This command turns string variables into numeric ones by putting an `n` (for "numeric") in the `op` column:

```{r str2num}
dat <- tibble(
  a1 = c("1", "0033"),
  b1 = c(3, 4),
)

cmd_block_tibble <- tibble::tribble(
  ~var, ~op,
  "a1", 'n', 
  "b1", '', 
)

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

## #AUTORECODE

### Example

This command autorecodes string variables by putting `a` (for "autorecode") in
the `op` column. It assigns labelled values (1, 2, 3, ...), ordered alphabetically:

```{r autorecode}
dat <- tibble(
  a1 = c("auto", "recode", "auto", "hello"),
)

cmd_block_tibble <- tibble::tribble(
  ~var, ~op,
  "a1", 'a', 
)

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

# `Label` sheets

Here, you can modify value labels or create new variable

```{r, echo=FALSE}
sheet_cat <- "Label"
```

## #SUMVAR

### Example

This command recodes variables specified in the `var` column (column **A**)
to new variables, by attaching a "k" prefix. It is very similar to the `#REC` command.

```{r sumvar}
dat <- tibble(
  a1 = haven::labelled(c(1, 2, 3, 4, 5), label = "original variable label", labels = purrr::set_names(1:5, LETTERS[1:5])),
)

cmd_block_tibble <- tibble::tribble(
  ~var, ~nv, ~sum_var_label, ~sum_var_value,	~sum_var_vallab,

  "a1", "1", 'new varlab', "1", "a & b", 
  "a1", "2", NA, "1", "" ,
  "a1", "3", NA, "2", "c" ,
  "a1", "4", NA, "3", "d & e",
  "a1", "5", NA, "3", "" ,
)

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

All values of `a1` (specified by `nv`) that are assigned to the same value
under `sum_var_value` are mapped onto that value (labelled by the string in the 
first row where this `sum_var_value` occurs). The assigned variable label of the 
summary variable is specified in the first row of `sum_var_label` (if empty the 
variable label of the original variable is copied).


## #NEWVALL

### Example

This command replaces value labels like `#AVALL`:

```{r newvall}
dat <- tibble(
  a1 = haven::labelled(c(1, 2, 3, 4, 5), label = "original variable label", labels = purrr::set_names(1:5, LETTERS[1:5])),
)

cmd_block_tibble <- tibble::tribble(
  ~var, ~nv, ~new_label,

  "a1", "1", 'new value label of value 1', 
  "a1", "2", NA, 
  "a1", "3", NA, 
  "a1", "4", NA, 
  "a1", "5", NA, 
)

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat, na_to_filter = FALSE)
```

# `#RECNA`

This command is executed per default (unless you set the parameter `na_to_filter`
to `FALSE` when creating the `Mapping` object with 

```{r, echo=TRUE, eval=FALSE}
Mapping$new(na_to_filter = FALSE)
```

for instance). It: 

* recodes missing values in all numeric 
variables in the dataset to 
`-2` (cf. parameter `miss_rec_val` in `gen_mapping_params()`), and 
* labels this value by 
`FILTER` (cf. parameter `miss_rec_lab` in `gen_mapping_params()`) in these 
variables

```{r recna}
sheet_cat <- "Free"
dat <- tibble(
  a = c(1, NA, 1), 
  b = haven::labelled(c(1, 2, NA), labels = c(a = 1, b = 2)),
  c = c(NA, "I'm a string", "variable"))

cmd_block_tibble <- tibble::tribble(
  ~X1, ~X2, ~X3,
  "#IF", "a == 1", "d = 2",
)

illustrate_cmd_block(dat, cmd_block_tibble, sheet_cat)
```


The missing values in the variables `a` & `b` were changed to `-2` and the value
label of `-2 = FILTER` was added. The variable `d` that was generated by the
`#COMP` command is not touched, as the `#RECNA` command is executed in the very
beginning of all commands in the mapping and thus only affects the numeric
variables present in the original dataset.


