---
title: "Translating the Excel mapping commands to R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Translating the Excel mapping commands to R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(datenanpassr)
```

When a `mapping` object is created,

```{r}
mapping_file <- system.file("extdata", "mapping.xlsx", package = "datenanpassr")
spss_file <- system.file("extdata", "fake_survey.sav", package = "datenanpassr")
m <- Mapping$new(spss_file, mapping_file)
```

the command blocks are translated to a list structure `m$cmd` containing intermediate 
results. 

## Excel sheets parsed

First is just a small data.frame containing the `sheet` name and its
`sheet_type`:

```{r}
m$cmd$sheet_cats
```

## Parse the sheets to R list objects

The sheets are parsed to a named list with R list objects in `m$cmd$sheet_data_raw`.
This is what the latter looks like for the  `"Free1"` sheet of the Excel file:

```{r}
m$cmd$sheet_data_raw$Free1
```

## Raw command table

The whole information in `m$cmd$sheet_data_raw` is then joint to a dataframe
```{r}
m$cmd$df_cmd_raw
```

## Generating `"command_block"`s

Each line in `m$cmd$df_cmd_raw` is translated to a `command_block()`. The list
of all lines is written to `m$cmd$command_blocks_raw`. See for example the 3rd 
element:
```{r}
m$cmd$command_blocks_raw[[3]]
```

## `"command_blocks"`

The list `m$cmd$command_blocks_raw` finally is translated to a 
`command_blocks()` object. These objects have their own printing method. See 
here for the 3 first elements:

```{r}
m$cmd$command_blocks[1:3]
```

But underneath, the structure is equal to `m$cmd$command_blocks_raw`
except an additional field `args`:
```{r}
names(m$cmd$command_blocks_raw[[3]])
names(m$cmd$command_blocks[[3]])
m$cmd$command_blocks[[3]]$args
```

## Overview command table

The whole information is put to `m$cmd_tbl`:

```{r}
m$cmd_tbl
```


## `args` overview table

Lets have a look at a table with an example of the types of the objects
in `args` for each of the `"command_block"` subclasses:

  * `x`, `y`: Character strings of variable names in the data set (or to be created).
  * `varlab` Character string containing a variable label
  * `v`: numeric() (or character()) vector; value(s) of a value label
  * `vallab`: character() vector; value label(s)
  * `id_list`: a list of ids in the data,
  * valid R character vectors of R expressions: 
    * `ex`: an expression, 
    * `ex_cond`: a conditonal expression, 
    * `ex_fun`: an existing function name
    * `filepath`: a filepath


<details>
  <summary>Click here to show intermediate steps to calculate `args_tbl2`! </summary>

```{r}
cdbs <- m$cmd_tbl
# Generate a `command_blocks` object with all unique types of
# `command_block`s  in the mapping file
(unique_cmd_tbl <- m$cmd_tbl[!duplicated(cdbs$action), ])
# Generate/print a list of all args sub-elements:

args_overview <- setNames(
  purrr::map(unique_cmd_tbl$command_blocks, "args"),
  unique_cmd_tbl$action
)
args_overview[1:3]


# Show a table of all `arg`s, and an example of their type & length
# for all command block types `cmd`:
args_tbl <- tibble::tibble(
  cmd = unique_cmd_tbl$action,
  example = args_overview) %>%
  tidyr::unnest_longer(example, indices_to = "args") %>%
  .[c(1, 3, 2)]

print(args_tbl, n = 1000)

args_tbl2 <- args_tbl %>% 
  dplyr::group_by(arg_class = args %>% 
    stringr::str_remove("\\d$") %>% 
    ifelse(. == "filepath", "ex_filepath", .) %>% 
    stringr::str_remove("(?<=ex)_.*") %>% 
    stringr::str_remove("s$")  
  ) %>% 
  dplyr::group_by(cmd) %>%
  dplyr::mutate(args = list(args))  %>%
  tidyr::pivot_wider(names_from = arg_class,
                     values_from = example,
                     values_fn = list) %>%
  dplyr::relocate(y, varlab, .after = x) 
args_tbl2
```

</details>

The following table shows an example of a subset of the columns arguments passed for each of the
command blocks:

```{r}
# This is a hack to print the named arguments. 
# If you use RStudio, the dataframe is best shown with:
# View(args_tbl2)
args_tbl2 %>% 
  dplyr::mutate(
    dplyr::across(
      x:id, 
      ~ paste(names(.x), .x, sep = "; ", collapse = "; ") %>%
        stringr::str_remove_all(
          'list\\(|\\)|c\\(|; NULL|"|^; '
        )
      )
    ) %>% 
  # show a subset of the columns:
    dplyr::select(-x, -id_list, -id, -args) %>% 
    print(n=222)
```



## Apply command blocks

These command blocks can then be applied to the data with 

```{r}
m$modify_data()
```

Underneath the hood, this function calls `apply_commands()` and passes it the arguments `args` (see above).

You can then access the data with
```{r}
m$dat_mod
```


<details>
  <summary>Click here to show whole list structure print output of `m$cmd`!</summary>
```{r}
m$cmd
```
  
</details>
