---
title: "Documentation of the Excel mapping command blocks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Documentation of the Excel mapping command blocks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(datenanpassr)
library(dplyr)
```

```{r child="../man/fragments/gt_table_stylers.Rmd", include=FALSE}
```

# Command block overview

In this document, we will show how the different command blocks in datenanpassr work. 
There are the following keywords that you can use to enter the command blocks in your Excel mapping file:
```{r}
command_block_classes %>% excelish()
```

*The column* `"command_block"` *shows the class, which determines how 
datenanpassr parses the command block* (with `parse_command_args()`) *and 
then applies it on the data* (with `apply_command()`).

# Command block examples

```{r}
command_block_doc_helper <- function(dat, cdb_args_raw) {
  cdb_class <- command_block_classes[command_block_classes$keyword == cdb_args_raw$X1[1], ][["command_block"]]
  class(cdb_args_raw) <- c(cdb_class, class(cdb_args_raw))
  cdb_args <- cdb_args_raw %>% parse_command_args()
  m0 <- Mapping$new()
  m0$dat_mod <- dat
  
  arg_list <- list(
    cdb = structure(
      list(),
      class = cdb_class
    ),
    mapping = m0
  ) %>% append(
    cdb_args
  )
  do.call(apply_command, arg_list)  
  
  list(
    cdb_class = cdb_class,
    cdb_args = cdb_args,
    result = m0$dat_mod
  )
  
}
```


## `#VARL`

```{r}
dat <- tibble(q1 = 4:6)

cdb_args_raw <- tibble(X1 = "#VARL", X2 = "q1", X3 = "assigned variable label")

res <- command_block_doc_helper(dat, cdb_args_raw)
```

```{r child="../man/fragments/command_block_example_template.Rmd"}
```

Actually, this command does not change the values, but it assigns the variable 
label specified in column **C** of the command to the variable specified in 
column **B**.

```{r}
# attributes(res$result$q1)
```


## `#VALL`

```{r}
dat <- tibble(q1 = c(1, 1, 2))

cdb_args_raw <- tibble::tribble(
  ~X1, ~X2,            ~X3,
  "#VALL", "q1", "yes/no question",
   NA,  "1",     "Yes",
   NA,  "2", "No"
  )

res <- command_block_doc_helper(dat, cdb_args_raw)
```

```{r child="../man/fragments/command_block_example_template.Rmd"}
```
This command also does not change the values. Just like `#VARL`, it assigns the variable label
specified in the first row of column **C** (if empty, it keeps the variable
label, if already existing) to the variable specified in the first row in column
**B**. It also assigns the value labels in column **C** to 
the values in column **B** in the following rows of the command.

```{r}
# attributes(res$result$q1)
```
