---
title: "Documentation of the Excel mapping command blocks"
output: 
  rmarkdown::html_vignette:
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Documentation of the Excel mapping command blocks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE
)
library(datenanpassr)
library(dplyr)
```

```{css css_def}
.parent {
  display: flex;
  height: 300px; 
}

.child {
  width: 100px;  
  height: 100px; 
  margin: auto;  
}


.flex-container {
  display: flex;
  align-items: center;
  flex-flow: row wrap;
  column-gap: 20px;
  row-gap: 10px;
}
```

```{r gt_style, child="../man/fragments/gt_table_stylers.Rmd", include=FALSE}
```
# Command block overview

In this document, we will show how the different command blocks in datenanpassr work. 
There are the following keywords that you can use to enter the command blocks in your Excel mapping file:
```{r cdb_classes}
command_block_classes %>% excelish()
```

*The column* `"command_block"` *shows the class, which determines how 
datenanpassr parses the command block* (with `parse_command_args()`) *and 
then applies it on the data* (with `apply_command()`).

# Command block examples

```{r fun_defs_illu, echo=FALSE}
process_cdb_raw <- function(dat, cdb_args_raw, sheet_cat, ...) {
  mapping_file <- gen_mapping_obj(cdb_args_raw, sheet_cat)
  Mapping$new(dat, mapping_file, ...)$modify_data()$dat_mod
}
gen_mapping_obj <- function(cdb_args_raw, sheet_cat) {
  if (sheet_cat == "Free") {
    cdb_args_raw <- bind_rows(
      tibble(
        X1 = character(),
        X2 = character(),
        X3 = character(),
        X4 = character(),
        X5 = character(),
      ),
      cdb_args_raw |> mutate(row = row_number())
    )
    mapping_file <- tibble::lst(!!sheet_cat := cdb_args_raw)
    return(mapping_file)
  }
}

tab_labels_wide <- function(df) {
  varl_table <- df |>
    datenanpassr:::tab_varlabs(remove_empty = FALSE) |> 
    tidyr::pivot_wider(names_from = var, values_from = varlab) |> 
    dplyr::mutate(val = "varlab", .before = 1)
  vall_table <- df |>
    datenanpassr:::tab_vallabs(remove_empty = FALSE) |>
    dplyr::rename(val = nv) |> 
    tidyr::pivot_wider(names_from = var, values_from = vallab) |> 
    dplyr::mutate(val = as.character(val))
  labs <- dplyr::bind_rows(varl_table, vall_table)
  if (nrow(labs) == 0) {
    labs <- labs |> dplyr::bind_rows(dplyr::tibble(val = NA_character_))
  }
  labs
}

join_labs <- function(df) {
  labs <- tab_labels_wide(df)
  df |>
    dplyr::mutate(
      dplyr::across(
        .cols = everything(),
        .fns = as.character
      ),
      val = NA_character_,
      .before = 1,
      type = "values"
    ) |> 
    dplyr::bind_rows(
      labs |> dplyr::mutate(type = "labels")
    )
}


print_cmd_exe <- function(dat, cdb_args_raw, sheet_cat, ...) {
  dat_mod <- process_cdb_raw(dat, cdb_args_raw, sheet_cat, ...)
  
  old_table <- join_labs(dat)
  new_table <- join_labs(dat_mod)
  
  join_cols <- dplyr::union(names(old_table), names(new_table))
  
  old_gt <- old_table |>
    gt::gt(
      groupname_col = "type"
    ) |> 
    gt::sub_missing(missing_text = "") |> 
    gt::cols_label(val = "") |> 
    color_gt_labels()
  new_gt <- new_table |>
    gt::gt(
      groupname_col = "type"
    ) |> 
    color_gt_labels()
  
  tibble::lst(
    old_gt,
    new_gt
  )

}
color_gt_labels <- function(gt_obj) {
  gt_obj |>
    gt::tab_style(
      style = list(
        gt::cell_fill(color = "#D4E6B4")
      ),
      locations = gt::cells_body(
        rows = val == "varlab")
    ) |> 
    gt::tab_style(
      style = list(
        gt::cell_fill(color = "#D7E9E9")
      ),
      locations = gt::cells_body(rows = type == "labels" & val != "varlab")
    ) |> 
    gt::tab_style(
      style = list(
        gt::cell_text(color = "#999999")
      ),
      locations = gt::cells_body(rows = type == "labels")
    ) |> 
    gt::sub_missing(missing_text = "") |> 
    gt::cols_label(val = "")
}

illustrate_cmd_block <- function(dat, cdb_args_raw, sheet_cat, ...) {
  cmd_block <- cdb_args_raw %>% purrr::set_names(LETTERS[seq_len(ncol(.))]) %>% excelish()
  
  gts <- print_cmd_exe(dat, cdb_args_raw, sheet_cat, ...)
  htmltools::div(
    class = "flex-container", 
    htmltools::div(style = "align-self: flex-start", gts$old_gt),
    htmltools::div(style = "font-size: 45px", "&"),
    htmltools::div(style = "flex-item", cmd_block),
    htmltools::div(style = "font-size: 45px", "=>"),
    htmltools::div(style = "align-self: flex-start", gts$new_gt),
  )
}
# This generates the following html:
# <div class="flex-container">
#   <div style="align-self: flex-start">`r gts$old_gt`</div>
#   <div style="font-size: 45px">&</div>
#   <div class="flex-item">`r cmd_block`</div>
#   <div style="font-size: 45px">=></div>
#   <div style="align-self: flex-start">`r gts$new_gt`</div>
# </div>

# if you want to see the result of this function interactively, pipe it in the following function:
# view_html <- function(x){
#   html <- paste(x, collapse = '\n')
#   tf <- tempfile(fileext = ".html")
#   writeLines(html, tf)
#   rstudioapi::viewer(tf)
# }

```

```{r, echo=FALSE}
sheet_cat <- "Free"
```


## `#VARL`

```{r varl}
dat <- tibble(q1 = 4:6)

cdb_args_raw <- tibble(X1 = "#VARL", X2 = "q1", X3 = "assigned variable label")

illustrate_cmd_block(dat, cdb_args_raw, sheet_cat, na_to_filter = FALSE)
```


Actually, this command does not change the values, but it assigns the variable 
label specified in column **C** of the command to the variable specified in 
column **B**.



## `#VALL`

Assign **val**ue **l**abels

```{r vall}
dat <- tibble(q1 = c(1, 1, 2))

cdb_args_raw <- tibble::tribble(
  ~X1, ~X2,            ~X3,
  "#VALL", "q1", "yes/no question",
   NA,  "1",     "Yes",
   NA,  "2", "No"
  )

illustrate_cmd_block(dat, cdb_args_raw, sheet_cat, na_to_filter = FALSE)
```

This command also does not change the values. Just like `#VARL`, it assigns the variable label
specified in the first row of column **C** (if empty, it keeps the variable
label, if already existing) to the variable specified in the first row in column
**B**. It also assigns the value labels in column **C** to 
the values in column **B** in the following rows of the command.


## `#AVALL`

**A**dd **val**ue **l**abels

```{r avall}
dat <- tibble(q1 = haven::labelled(c(1, 2, 3), c(a = 1)))

cdb_args_raw <- tibble::tribble(
  ~X1, ~X2,            ~X3,
  "#AVALL", "q1", NA_character_,
   NA,  "2",     "b",
   NA,  "3", "c"
  )

illustrate_cmd_block(dat, cdb_args_raw, sheet_cat, na_to_filter = FALSE)
```

`#AVALL` does the same as `#VALL`, except that it keeps the value values of the 
original variable, if they are not overwritten.

## `#RMVAL`

**R**e**m**ove **val**ues and their labels

```{r rmval}
dat <- tibble(q1 = haven::labelled(c(1, 2, 3), c(a = 1, b = 2, c = 3), "abc"))

cdb_args_raw <- tibble::tribble(
  ~X1, ~X2,            ~X3, ~X4,
  "#RMVAL", "q1", "q2", "def",
   NA,  "2",     "", NA_character_,
   NA,  "3", "", NA_character_,
  )

illustrate_cmd_block(dat, cdb_args_raw, sheet_cat, na_to_filter = FALSE)
```

Here, the `#RMVAL` command makes a copy of `q1`, named `q2`, and removes the 
specified values.
(In the first line of the command block, column **B** specifies the variable to be recoded, 
and in column **C** the name of the recoded variable. Column **D** assigns the new 
variable label.) In the following rows you specify the values and their labels 
to be removed in column **B**.

## `#COMP`

### Example

```{r comp1}
dat <- tibble(q1 = 1:3)

cdb_args_raw <- tibble::tribble(
  ~X1, ~X2,            ~X3,
  "#COMP", "q2", "5",
  )

illustrate_cmd_block(dat, cdb_args_raw, sheet_cat, na_to_filter = FALSE)
```
This command **comp**utes a new variable (specified in column **B**) with the value 
in column **C**.

### Example

You can also assign a variable to the values of the new variable. In this case,
the labels of the original variable are also copied:
```{r comp2}
dat <- tibble(q1 = haven::labelled(c(1, 2, 3), c(a = 1, b = 2, c = 3), "abc"))

cdb_args_raw <- tibble::tribble(
  ~X1, ~X2,            ~X3,
  "#COMP", "q2", "q1",
  )

illustrate_cmd_block(dat, cdb_args_raw, sheet_cat, na_to_filter = FALSE)
```

## `#IF`

### Example

```{r if1}
dat <- tibble(q1 = 1:3)

cdb_args_raw <- tibble::tribble(
  ~X1, ~X2,            ~X3,
  "#IF", "q1 >= 2", "q2 = 5",
  )

illustrate_cmd_block(dat, cdb_args_raw, sheet_cat, na_to_filter = FALSE)
```
This command takes a condition (more precisely a logical vector of length of the 
variables in the data) in column **B** and an assignment in column **C** (of the form 
`<variable name> = <assigned values>`). 
When the condition is `TRUE`, the expression (`<assigned values>`) is assigned 
to `<variable name>`.
In this case, the value 5 is assigned to `q2` where `q1` is greater or equal to 2.
In this case `q2` didn't exist in the data before, and thus is initialized to
`NA` before.

### Example

If the variable existed before, the values where the condition evaluates to `FALSE`
are kept, as can be seen in the next example:

```{r if2}
dat <- tibble(q1 = 1:3, q2 = c(7, 8, 9))

cdb_args_raw <- tibble::tribble(
  ~X1, ~X2,            ~X3,
  "#IF", "q1 >= 2", "q2 = 5",
  )

illustrate_cmd_block(dat, cdb_args_raw, sheet_cat, na_to_filter = FALSE)
```

Here the value of 7 in the first row is kept from the original data.

### Example

The `#IF` command preserves labels, if an already existing variable is modified:

```{r if3}
dat <- tibble(q1 = haven::labelled(c(1, 2, 3), c(a = 1, b = 2, c = 3), "abc"))

cdb_args_raw <- tibble::tribble(
  ~X1, ~X2,            ~X3,
  "#IF", "q1 >= 2", "q1 = 5",
  )

illustrate_cmd_block(dat, cdb_args_raw, sheet_cat, na_to_filter = FALSE)
```

## #REC

### Example

```{r rec1}
dat <- tibble(q1 = haven::labelled(c(1, 2, 3, 4, 5, 99), c(a = 1, b = 2, c = 3, d = 4, e = 5, "No answer" = 99), "abcde"))

cdb_args_raw <- tibble::tribble(
  ~X1, ~X2,            ~X3, ~X4, ~X5,
  "#REC", "q1", "kq1", "fgh", NA_character_,
  "", "1", "2", "1", "a & b",
  "", "3", "",  "2", "c",
  "", "4", "5", "3", "d & e",
  )

illustrate_cmd_block(dat, cdb_args_raw, sheet_cat, na_to_filter = FALSE)
```

The `#REC` command **rec**odes variables. In this example `q1` is recoded to `q2`.
(In the first line of the command block, column **B** specifies the variable to be recoded, 
and in column **C** the name of the recoded variable. Column **D** assigns the new 
variable label.) In the following rows you specify the intervals to be recoded 
in column **B** and **C**, column **D** specifies the new value and column **E**
its label. So for instance, the second line specifies, that all values in `q1` 
between 1 and 2 are recoded to 1 in `q2` with the value label `a & b`.

### Example

If you don't specify the variable label (column **D**, first row), the one of 
the original variable is copied:

```{r rec2}
dat <- tibble(q1 = haven::labelled(c(1, 2, 3), c(a = 1, b = 2, c = 3), "abc"))

cdb_args_raw <- tibble::tribble(
  ~X1, ~X2,            ~X3, ~X4, ~X5,
  "#REC", "q1", "kq1", NA_character_, NA_character_,
  "", "1", "2", "1", "a & b",
  )

illustrate_cmd_block(dat, cdb_args_raw, sheet_cat, na_to_filter = FALSE)
```

### Example

If the name of the recoded variable is the same as the original variable, the 
values that are not recoded are kept:
```{r rec3}
dat <- tibble(q1 = haven::labelled(c(1, 2, 3), c(a = 1, b = 2, c = 3), "abc"))

cdb_args_raw <- tibble::tribble(
  ~X1, ~X2,            ~X3, ~X4, ~X5,
  "#REC", "q1", "q1", NA_character_, NA_character_,
  "", "1", "2", "0", "a & b",
  )

illustrate_cmd_block(dat, cdb_args_raw, sheet_cat, na_to_filter = FALSE)
```

## #KG

### Example

```{r kg1}
dat <- tibble(
  q1 = haven::labelled(
    c(1, 2, 1, 2, 1, 2), 
    c(good = 1, bad = 2),
    "How are you?"
  ),
  q2 = haven::labelled(
    c(1, 2, 3, 1, 2, 3), 
    c(UK = 1, US = 2, FR = 3, DE = 4),
    "Country"
  ),
)

cdb_args_raw <- tibble::tribble(
  ~X1, ~X2, ~X3, 
  "#KG", "q1", "q2", 
  )

illustrate_cmd_block(dat, cdb_args_raw, sheet_cat, na_to_filter = FALSE)
```

The `#KG` command is useful to calculate cross-tabulations for subsets. It 
computes new variables for all the value labels of the variable `q2` (in column **C**).
The value labels are copied from `q1` (the variable in column **B**). The new variable 
labels are a combination of the value labels of `q2` and the variable label of `q1`.
The values are copied of `q1` if `q2` takes the value specified in the variable 
label and are `NA` otherwise.
