---
title: "Translating the Excel mapping commands to R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Translating the Excel mapping commands to R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(datenanpassr)
```

When a `mapping` object is created,

```{r}
mapping_file <- system.file("extdata", "mapping.xlsx", package = "datenanpassr")
spss_file <- system.file("extdata", "fake_survey.sav", package = "datenanpassr")
m <- Mapping$new(spss_file, mapping_file)
```

the command blocks are translated to a list structure `m$cmd` containing intermediate 
results. 

## Excel sheets parsed

First is just a small data.frame containing the `sheet` name and its
`sheet_type`:

```{r}
m$cmd$sheet_cats
```

## Parse the sheets to R list objects

The sheets are parsed to a named list with R list objects in `m$cmd$sheet_data_raw`.
This is what the latter looks like for the  `"Free1"` sheet of the Excel file:

```{r}
m$cmd$sheet_data_raw$Free1
```

## Command table

The whole information in `m$cmd$sheet_data_raw` is then joint to a dataframe
```{r}
m$cmd$df_cmd_raw
```

## Generating `"command_block"`s

Each line in `m$cmd$df_cmd_raw` is translated to a `command_block()`. The list
of all lines is written to `m$cmd$command_blocks_raw`. See for example the 3rd 
element:
```{r}
m$cmd$command_blocks_raw[[3]]
```

## `"command_blocks"`

The list `m$cmd$command_blocks_raw` finally is translated to a 
`command_blocks()` object. These objects have their own printing method. See 
here for the 3 first elements:

```{r}
m$cmd$command_blocks[1:3]
```

But underneath, the structure is equal to `m$cmd$command_blocks_raw`, with an
additional field `args` added:
```{r}
m$cmd$command_blocks[[3]]$args
```

## Apply command blocks

These command blocks can then be applied to the data with 

```{r}
m$modify_data()
```

Underneath the hood, this function calls `apply_commands()` and passes it the arguments `args` (see above).

You can access the data with
```{r}
m$dat_mod
```


<details>
  <summary>Click here to show whole list structure print output!</summary>
```{r}
m$cmd
```
  
</details>
