---
title: "debugging"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{debugging}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Debugging

In order to debug the command blocks of a mapping you can set the option
`debug = TRUE`.

```{r}
library(datenanpassr)
mapping_file <- system.file("extdata", "mapping.xlsx", package = "datenanpassr")
m <- Mapping$new(fake_survey, mapping_file, debug = TRUE)
```

When the execution of a code block runs into an error, this then jumps into the code.
We'll sneak an error in one of the `"command_block"`s:
```{r}
m$cmd_tbl$command_blocks[[52]]$args$ex_cond <- "q1 ==(*%$@ 1 |} q3 == 2"
```

Now, when we execute the modifications, the debugger stops at the `browser()`.
```{r}
m$modify_data()
```

and you will see the command prompt of the debugger:

`Browse[1]> `

There, you can move into the function that produces the error by pressing `Shift` + `F5`.
