---
title: "debugging"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{debugging}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

In order to debug the command blocks of a mapping you can set the option
`debug == TRUE`.

```{r}
library(datenanpassr)
mapping_file <- system.file("extdata", "mapping.xlsx", package = "datenanpassr")
m <- Mapping$new(fake_survey, mapping_file, debug = TRUE)
```

When the execution of a code block runs into an error, this then jumps into the code.

```{r}
m$modify_data()
```

When the debugger stops at the `browser()` you can also study the data with other
software depending on the filetype:

* `"sav"` : SPSS
* `"dta"` : Stata
* `"xlsx"`: Excel file with the labels removed
* `"Rmd"` : Setup a python session in Rmarkdown

When the debugger stops, `self` represents the current `Mapping` object. 
The following code would then fire up all these programs with the current data:
```{r}
self$save(c("a.sav", "a.dta", "a.Rmd", "a.xlsx"), show = TRUE)
```

You can also only show part of the files generated by passing a vector of the 
same length as the `path` argument.
```{r}
self$save(path = c("a.sav", "a.dta", "a.Rmd", "a.xlsx"), show = c(FALSE, FALSE, FALSE, TRUE))
```

