---
title: "Included data sets and files in the package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Included data sets and files in the package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(datenanpassr)
library(haven)
```

### Included labelled data

The package has an included dataset `fake_survey` with made-up data of a survey. The data is also included as an SPSS sav file [fake_survey.sav](../inst/extdata/fake_survey.sav). These are generated by [this Rmd file](../devel/make_fake_data.md). After installing the package, here is how you can access these data sets: 


```{r example}
fake_survey
path <- system.file("extdata", "fake_survey.sav", package = "datenanpassr")
read_sav(path)
```

### Mapping file

An example of commands that can be entered in the mapping file created by `mapp_create()` can be found [here](../inst/extdata/mapping.xlsx)


### Included verbatim file 

The assignment of labelled codes to open answers can be done in the  [verbatim file](../inst/extdata/Verbatims_fake_survey.xlsx).

