---
title: "OOP_structure"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup}
library(datenanpassr)
```

### Excel files

The package is works with Excel files, such as the one included in the package:
```{r}
mapping_file <- system.file("extdata", "mapping.xlsx", package = "datenanpassr")

```

These can be transformed to objects of class "mapping" with 
```{r}
(mapping <- new_mapping(mapping_file))
```

### Command table

The field `df_cmd` is initialized as `NULL` but can be computed with
```{r}
(df_cmd <- mapp_cmd_table(mapping))
```

Under the hood, this transforms the mapping structure, by calling 

```{r}
mapping_cmd <- mapp_cmd_table_(mapping)
```

This creates an object of subclass `"cmd_table"`, inheriting from the "mapping"
class and with object returned by `mapp_cmd_table()` added.

### Data adaptations 

In order to apply the commands from the Excel file to the dataset `fake_survey`
(included in the package), you can call the function

```{r}
(mapping_mod <- mapp_xl_to_data_(df = fake_survey, mapping = mapping_cmd))
```

This adds the subclass `"data_mod"` to the mapping object.

To directly extract the modified data set from this structure:

```{r}
(df_mod <- mapp_xl_to_data(fake_survey, mapping_cmd))
```



### Sequence of methods

As above, the execution of the shown methods can be done beforehand and then passed into 
the following methods.

Alternatively, you can also directly call the functions `mapp_xl_to_data_()` and `mapp_cmd_table_()` (as well as their counter parts without the underscore `_` suffix)
with mapping file paths (instead of objects of class `"cmd_table"` and `"mapping"`, respectively):

```{r, eval=FALSE}
mapping_cmd <- mapp_cmd_table_(mapping_file)
mapping_mod <- mapp_xl_to_data_(df = fake_survey, mapping = mapping_file)
```

Under the hood, this is done with the `as_mapping()` and `as_cmd_table()` functions.


### Summary

* An object of the main class `"mapping"` provided by this package can be built
by calling the low-level constructor `new_mapping()`. It needs a file path of a
valid mapping file and returns a list structure.
* The constructor `mapp_cmd_table_()` takes a "mapping" object,
adds the element `df_cmd` (resulting of the mapping file) to the list structure
and returns an object of the subclass `"cmd_table"`.
* `mapp_xl_to_data_()` takes an object of class `"cmd_table"` and additionally
needs as an input argument the data set `df` on which the commands in `df_cmd`
should be applied on. It then computes `df_mod` (resulting when the commands in
`df_cmd` are applied to `df`), adds it to the list structure . Finally the list
structure is returned with the added subclass `"data_mod"`:

```{r}
class(mapping)
class(mapping_cmd)
class(mapping_mod)
```



