---
title: "datenanpassr Package Report, by pkgnet"
output: 
    rmarkdown::html_vignette:
        toc: true
vignette: >
    %\VignetteIndexEntry{Package Report, by pkgnet}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

# My customized package functions dependency graph

* The graph shows the functions of the package (nodes)
* Scroll to zoom in/out
* Drag n drop to move focus/node
* Exported functions of the package are written in bigger font
* Colors and shapes denote the files where the functions are defined
* Hover over/click on nodes to see the local hierarchical dependency structure of the functions (the other functions it calls/is called by)
* Select by group to see the functions of the different files of the package
* Double-click on nodes to collapse (and then re-explode) its child nodes (functions the function calls)

```{r message=FALSE, include=FALSE}
library(dplyr)
library(purrr)
library(tidygraph)

fun_deps <- pkgnet::FunctionReporter$new()

fun_deps$set_package(pkg_name = "datenanpassr")

fun_dep_graph <- fun_deps$pkg_graph$igraph %>% tidygraph::as_tbl_graph()

# from here: https://stackoverflow.com/questions/20259681/how-to-list-all-the-functions-signatures-in-an-r-file/20260081#20260081

is_function = function (expr) {
  if (! is_assign(expr))
    return(FALSE)
  value = expr[[3]]
  is.call(value) && as.character(value[[1]]) == 'function'
}

function_name = function (expr)
  as.character(expr[[2]])

is_assign = function (expr)
  is.call(expr) && as.character(expr[[1]]) %in% c('=', '<-', 'assign')

l <- fs::dir_info(here::here("R")) %>% pull(path) %>% set_names() %>% map(parse)
functions = l %>% map(~Filter(is_function, .x))

function_names = functions %>% map(~unlist(Map(function_name, .x)))

df_fun_file <- functions %>% map(~unlist(Map(function_name, .x))) %>% map_dfr(as_tibble, .id = "file")

df_node_files <- df_fun_file %>% rename(name = value) %>% distinct(name, .keep_all = TRUE)

graph_obj <- fun_dep_graph %>%
  # as_tibble() %>%
  left_join(df_node_files) %>%
  left_join(fun_deps$nodes %>% as_tibble() %>% rename(name = node)) %>%
  mutate(ex_name = ifelse(isExported, name, NA))
# library(ggraph)
# graph_obj %>%
#   ggraph(
#     # layout = "tree",
#     # circular = TRUE
#   ) +
#   geom_edge_link() +
#   ggforce::geom_mark_hull(aes(x = x, y = y, fill = file, label = file)) +
#   geom_node_point(aes(color = file, size = 3, shape = file), stroke = 2) +
#   scale_shape_manual(values=seq(0,length(function_names))) +
#   geom_node_label(aes(color = file, label = ex_name, size = 3), repel = TRUE, max.overlaps = 50)


# visnetwork customization ------------------------------------------------

all_shapes <- c("ellipse",
                # "circle", "database",
                "box",
                # "text",
                # "image", "circularImage",
                "diamond",
                "dot",
                "star", "triangle", "triangleDown", "square"#, "icon"
                )

df_nodes <- graph_obj %>%
  as_tibble() %>%
  transmute(
    id = forcats::as_factor(name) %>% as.numeric(),
    label = name,
    shape = factor(file) %>% as.numeric() %>% {. %% length(all_shapes)} %>% {. + 1} %>% all_shapes[.],
    # shape = all_shapes[shape],
    color = rainbow(length(function_names))[factor(file) %>% as.numeric()],
    value = ifelse(isExported, 20, 10),
    font.size = ifelse(isExported, 20, 10),
    group = file
  )

df_edges <- graph_obj %E>%
  as_tibble() %>%
  mutate(arrows = "to")

v <- visNetwork::visNetwork(df_nodes, df_edges, height = "800px", width = "150%") %>%
  visNetwork::visOptions(
    highlightNearest = list(enabled = TRUE, degree = 10, hover = TRUE, algorithm = "hierarchical"),
    nodesIdSelection = TRUE,
    collapse = TRUE,
    selectedBy = "group"
  )

```

```{r}
v
```

# pkgnet


This report on **datenanpassr** is generated by [**pkgnet**](https://uptakeopensource.github.io/pkgnet/), an R package for analyzing other R packages through the lens of [network theory](https://en.wikipedia.org/wiki/Network_theory). 

```{r, include=FALSE}
library(pkgnet)

knitr::opts_chunk$set(
    echo = FALSE, 
    warning=FALSE, 
    out.width='100%'
)
pkgnet:::silence_logger()

reporters <- list(DependencyReporter$new(), FunctionReporter$new())

reporters <- lapply(
          X = reporters
          , FUN = function(reporter){
              reporter$set_package(pkg_name = "datenanpassr")
              return(reporter)
          }
      )

```


```{r warning=FALSE}
reportSections <- lapply(reporters, function(reporter) {
  report_env <- list2env(list(reporter = reporter))
  knitr::knit_child(
    reporter$report_markdown_path
    , envir = report_env
  )
})

```

```{r results="asis"}
cat(paste0(paste(reportSections, collapse = '\n')))
```

---

```{r results="asis"}
cat(sprintf("<sup>This report built with **pkgnet v%s**.</sup>", packageVersion('pkgnet')))
```

<sup>[**pkgnet**](https://uptakeopensource.github.io/pkgnet/) is an open-source R package, copyright &copy; 2017-2019 Uptake, made available under the 3-Clause BSD License.</sup>

```{r echo = FALSE}
pkgnet:::unsilence_logger()
```


