---
title: "benchmark_data_manipulations"
author: "Urs Wilke"
date: "3/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE)
```

```{r libs, message=FALSE, echo=FALSE}
# library(dtplyr)
# library(rlang)
library(tidyverse)
# `%>%` <- magrittr::`%>%`
```

This report compares execution times of different approaches for consecutive data manipulations of a dataframe, like it is done in `mapp_xl_to_data()`.

4 approaches are compared:

* manipulations with a single `mutate()`
* manipulations with a data.table (using dtplyr)
* manipulations with a a recursive `mutate()` (using `purrr::reduce()`)
* manipulations with a a recursive `mutate()` (using `purrr::reduce()`) AND restricting the manipulated dataframe on the columns needed (like it is done until now...)

It is measured how much time it takes to manipulate this data.frame:

```{r explain}
df
```

with these manipulations (Example for 10 manipulations...):

```{r manip}
idx <- 1:10
l <- list(var = paste0("v", idx + 1), val = rlang::parse_exprs(paste0("v", idx, " + 1")))
# created column names
l$var
# manipulation operations:
l$val
```



resulting in this dataframe:

```{r dfmod}
df %>% mutate(!!!set_names(l$val, l$var))
```

```{r current}
cmd_comp <- function(df, new_var, new_val) {
  if (!new_var %in% names(df)) {
    df[new_var] <- NA_real_
  }
  varlab <- labelled::var_label(df[[new_var]])
  vallabs <- labelled::val_labels(df[[new_var]])

  manipulated_vars <- datenanpassr:::get_df_vars_of_expr_string(paste(new_val), names(df)) %>%
    c(new_var) %>% unique()

  new_val <- rlang::parse_expr(new_val)
  df[manipulated_vars] <-
    df[manipulated_vars] %>% dplyr::mutate(!!rlang::sym(new_var) := !!new_val)
  df
}
```


```{r funs}
f <- function(df, var, val) {
  dplyr::mutate(df, !!var := !!val)
}
```


```{r bench_fun}
bench_ops_no <- function(id_max) {
  idx <- 1:id_max
  l <- list(var = paste0("v", idx + 1), val = rlang::parse_exprs(paste0("v", idx, " + 1")))
  l2 <- list(var = paste0("v", idx + 1), val = paste0("v", idx, " + 1"))
  # df_mod <- purrr::reduce2(l$var, l$val, f, .init = df) %>% tibble::as_tibble()
  # df_mod <- purrr::reduce2(l$var, l$val, f, .init = dtplyr::lazy_dt(df)) %>% tibble::as_tibble()
  microbenchmark::microbenchmark(
    single = df %>% mutate(!!!set_names(l$val, l$var)),
    reduce = df_mod <- purrr::reduce2(l$var, l$val, f, .init = df),
    reduce_restr = df_mod <- purrr::reduce2(l2$var, l2$val, cmd_comp, .init = df),
    data.table = df_mod <- purrr::reduce2(l$var, l$val, f, .init = dtplyr::lazy_dt(df)) %>% tibble::as_tibble()
  )
}

```


```{r benchmarking}

df <- data.frame(v1 = 1:10)

exponential_ops_no_series <- 2^(1:7) %>%
  floor() %>%
  set_names()
benches <- exponential_ops_no_series %>%
  purrr::map(bench_ops_no)
```


```{r benchmark_wrangling}
bench_df <- benches %>%
  map(
    ~ .x %>%
      group_by(expr) %>%
      summarise(t = mean(time))
  ) %>%
  imap_dfr(~.x, .id = "n_manips") %>%
  mutate(n_manips = as.numeric(n_manips))
```


```{r benchmark_plot}
p1 <- bench_df %>%
  ggplot(aes(n_manips, t / 1e9, color = expr)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Mean execution time",
    x = "Number of manipulations",
    y = "execution time (s)"
  )
plotly::ggplotly(p1)
```


```{r benchmark_plot_log}
p2 <- p1 +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    title = "Mean execution time (log scale)",
    x = "Number of manipulations",
    y = "execution time (s)"
  )
plotly::ggplotly(p2)
```


```{r benchmark_plot_compare}
t_s <- bench_df %>% filter(expr == "single") %>% pull(t)

p3 <- bench_df %>%
  group_by(expr) %>%
  mutate(t_rel = t / t_s) %>%
  ggplot(aes(n_manips, t_rel, color = expr)) +
  geom_point() +
  geom_line() +
  scale_x_log10(
    # breaks = scales::trans_breaks("log2", function(x) 2^x),
    # labels = scales::trans_format("log2", scales::math_format(2^.x))
  ) +
  scale_y_log10(
    # breaks = scales::trans_breaks("log10", function(x) 10^x),
    # labels = scales::trans_format("log10", scales::math_format(10^.x))
  )  +
  labs(
    title = "Ratio of mean execution time compared to fastest method",
    x = "Number of manipulations",
    y = "ratio of execution times"
  )

plotly::ggplotly(p3)


```

